### 2018-11-22 paper.tex

\documentclass[runningheads]{llncs}
\usepackage{bm}
\usepackage{CJK}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{tikz}
\usetikzlibrary{fit,positioning}

\begin{document}
\title{Explain urban regional function}
\author{Submission}

\maketitle

\begin{abstract}
Functional region identification is a critical step towards Urban computing.
The state-of-the-art Dirichlet Multinomial Regression(DMR) has identify regions with different function based on topic cluster modeling.
Due to the ingenious structure inside the complex model, its result is beyond the direct comprehension of humans.
How to give them reliable explanations are the main objectives of this work.
To improve the persuasiveness of its result, we proposed a post hoc framework to make explanation for it.
Optimal latent factor model was applied as explanation after the identification.
We use matrix factorization to get latent factors and correlate latent factors with urban feature.
The selected urban feature can be labeled as the explanation.
Our framework can give an strong evidence to the identification and enhance the satisfaction of users and system designers.

(15--250 words.)

\keywords{Explanation  \and Urban computing \and Online content \and Sentiment analysis \and Latent Factor \and Post hoc.}
\end{abstract}
%

\section{Introduction}
%urban computing
Urban computing, defined in~\cite{Zheng2014UrbanConcepts} as a process of acquisition, integration and analysis of big and heterogeneous data generated by diverse sources in urban spaces.%definition
One of the most important data source is human mobility data, e.g. pick-ups and drop-offs of taxicabs.
Such mobility data can contribute to improve living quality of residents, e.g. optimizing urban planning~\cite{Zheng2011taxicabs}, easing traffic congestion~\cite{Meng2017Traffic}, decreasing energy consumption ~\cite{Zhang2015Fuel}, reducing air pollution~\cite{Shang2014Pollution} and so on.
Therefore, mining and modeling mobility data has attracted attentions from both academia and industry.

%region function
Identifying functional regions has been a critical step towards efficient government administration and policy making for decades.
In early work, a functional region was defined as a geographical region where the majority of local population recruit and are employed within the region \cite{Ball1980Definition}.
Due to the availability of large-scale mobility data, an increasing amount of data-driven approaches~\cite{Farmer2009Overview} have been proposed.
Most of them use clustering methods on commuting data, such as origin-destination pairs of labour market data ~\cite{Karlsson2006FunctionalRegionSummary} while others use remote-sensor image data~\cite{Vatsavai2011Remote}.
Later work focuses on city-level functional region identification, where a functional region is a region in a city which supports different needs of people’s urban lives~\cite{Yuan2012FunctionRegion,Yuan2015FunctionRegion}.
To obtain accurate identification, recent research tends to use complex models for various form of data,
e.g. latent factor models based on Dirichlet Multinomial Regression are applied on mobility data of taxicabs and points of interests (POIs)~\cite{Yuan2012FunctionRegion,Yuan2015FunctionRegion}, clustering algorithm based on the 'modularity function' is applied on telecommunication~\cite{Newman2004ModularityFunction,Ratti2010Telecom},unsupervised semantic labeling framework based on the Latent Dirichlet Allocation is applied on remote-sensor data~\cite{Vatsavai2010Remote}. %complex model

\textbf{Motivation.}
%explainability
A severe drawback of existing research is the lack of explanation for the identifications of functional regions. Though existing applications~\cite{Yuan2015FunctionRegion} have achieved high identification accuracy,the complex nature of these models weakens the interpretability to end users and system designers.%The lack of explanation of the identification xxx
There is an emerging trend in studying explainable AI~\cite{Preece2018ExplainAI}.
In geographical systems, explainable system is still in its initial stage~\cite{Korpan2017navigation,Jose2018cognitive}.
To the best of our knowledge, none of the previous urban functional identification provide explanation for the results of region segmentation and functional labels.

%explain implementation, we need to incorporate activity and online sentiment
\textbf{Framework.}
In this paper,to make the results more convincing, we build a system that provides explanations for functional region clustering results.
The system is built upon two data sets. One is mobility data including more than 3 million pick-up and drop-off coordinates and timestamps,which provide fuzzy flow increment of each region units and each time bins.
The increment, also called moving patterns, are the first input of topic-model-based DMR models as well as POIs.
Dirichlet Multinomial Regression(DMR) identify the segmented region units as different functions.
Meanwhile, comment data of shop show the sentiment relavitity between urban feature and the activity that shop related to.
We pick labels with highest information entropy as urban features from all extracted features.
Finally, we optimal latent factor with temporal dimension to get a location-feature matrix.

As shown in Fig.\ref{explanation-example}, our system not only segments regions of a city based on the functionals, but also delivers explanations of the  functionals. To be specific, we first associate each functional with an urban activity, next we extract the most representative urban feature for each activity, finally we highlight the sentiment towards typical urban features for each functional region.  For example,Fig.\ref{explanation-example} given a map of city A, reg region I corresponds to residential region while blue region II corresponds to commercial region. We found that people are most interested on traffic for a commercial region. And a residential region usually has sentiment positive on vanishing noise.

\begin{figure}
\centering
\subfigure[An example of our identification result]{
\label{real-map} %% label for first subfigure
\includegraphics[width=1.5in]{example-net.png}}
\subfigure[moving patterns]{
\label{patterns} %% label for second subfigure
\includegraphics[width=1.4in]{example-line.png}}
\subfigure[POIs in different functional region]{
\label{sentiment} %% label for second subfigure
\includegraphics[width=1.4in]{example-sentiment.png}}
\caption{The explanation provided for functional region identification.
Region I in Fig.\ref{real-map} is labeled as residential region while II is commercial region.
Fig.\ref{patterns} revealed differences of the mobility patterns between residential region and commercial region,workdays and weekends.
Fig.\ref{POI} shows the distribution of POIs within the two regions,e.g. there is more shops in commercial region.
And from fig.\ref{sentiment} we find that people have a more interest in a positive sentiment in traffic for commercial and pay more attention to noise for residential region.}
\label{explanation-example} %% label for entire figure
\end{figure}

\textbf{Challenge.}
In order to solve the problem, we face two challenges.

The first one of it is to extract the urban feature from comment data.
Urban feature is a set of characters with special geographical attributes, e.g., clothes, school, bars and so on.
Most of our content data are comments of shops with the shop address.The feature extracted from it is more related to shop feature than urban feature.
Therefore, existing sentiment extraction model ~\cite{Lu2011Label} can not be used directly in our models.%urban feature
In order to get the location information from these content data, all features are extracted as~\cite{Zhang2015Sentires}.
We select typical features with a high frequency and high information entropy from it.

And the second challenge is how to explain the features.
If the extracted urban features used for explanations can not correspond to the labeled activity used to identify functional regions, the explanations will not have a strong explanation power for the identification, or even will not be generated.
To improve the explanation power of our framework, we extract related urban feature and find the corresponding sentiment score between activity and urban feature from contents.
With the probability that region units belong to which function generated by DMR, we can applied matrix factorization to find the relativity between region units and latent factor, i,e urban feature.
The urban feature with highest relativity make an explanation for why corresponding region units are identified as such function.

\textbf{Contribution.}
Our paper has an outstanding result in following several aspect:
\begin{itemize}
  \item To the best of our knowledge,we are the first to give explanation to urban functional region identification , which could improve the trustiness and satisfaction of users and urban system designers.
  \item After identify different functions, We use the contents with sentiment to find the typical urban feature within a region and make explanation.
  \item The explanation is given by latent factor model. We appplied matrix factorization to get the correlation coefficients between single region units and its features.
\end{itemize}

\textbf{Structure.}
The remainder of the paper is organized as follows.
We give an overview for related work in Section 2.
In Section 3, we give an overview of our framework.
Section 4 shows the process of functional region identification as well as urban feature extraction.
And Section 5 introduces the post hoc explanation model.
We test the framework on our datasets and presents the superior explanation result in Section 6
Finally, we made a conclusion and looked forward to our future work in Section 7.

\section{Related Work}
Our work is build upon taxicabs trajectory and comment text.
With the data source, two lines of work are related to this paper: urban computing and online sentiment analysis .

%related work section name, find another suitable name for mobility mining
\subsection{Urban Computing}

Urban computing~\cite{Zheng2014UrbanConcepts} tackles the major issues that cities face by analyzing human mobility collected from different sensors.
Major sources of human mobility data are check-ins in POI~\cite{Gallegos16happier}, pick-up and drop-off behavior of taxicabs~\cite{Ge2011TaxiBusiness,Meng2017Traffic} in different locations and  trajectories.

%extract the different among these data
The simplest form of mobility data is \textbf{check-ins data}, which are collected from locating sharing services. Check-ins data usually includes a set of point revealing users' current location.
Statistical association analysis is commonly conducted on check-ins. For example, radius of gyration is measured in~\cite{Cheng2011Check-in}, which is extended by combining with lexicon into demographics~\cite{Gallegos16happier}.
A few recent work adopts model based approaches, e.g. context-aware tensor factorization that take account of contextual factors that influence consumers’ refueling decision~\cite{Zhang2015Fuel}.

The second form of mobility data is \textbf{origin-destination pairs}, i.e. a pick-up point and a drop-off point of a taxi trajectory.
Clustering methods including Newman modularity cluster algorithm~\cite{Newman2004ModularityFunction} are applied in functional region identification~\cite{Ratti2010Telecom}.
Recently, latent factor models are proposed to treat a region as document and infer functional-specific ~\cite{Yuan2012FunctionRegion}.%
Latent activity is imported in topic model in~\cite{Yuan2015FunctionRegion} to define the specific functions of different regions.

Alternatives for origin-destination pairs are \textbf{trajectory data}.
Applications on trajectory data include recommendation, e.g. to recommend more suitable place to drivers~\cite{Ge2011TaxiBusiness,Garg2018Route} and traffic planning, e.g. to find reachable region within a given temporal period~\cite{Wu2017Reachable}, %find a fast route
or to predict travel time~\cite{Wang2018Deep}.

%tasks
However, existing urban computing systems extensively rely on complex machine learning algorithms hence they act as blank-boxes for end users.
The lack of explanation weakens the persuasiveness and trustworthiness of the system for users.
Our work is to make up for this drawback by providing intuitive explanations of the results for users or system designers

\subsection{Geographical Analysis of Online Sentiment}

Recently, an emerging research interest is witnessed in exploring the geographical factors that affect online sentiment.
Empirical studies have been conducted on large-scale human mobility data, such as check-in ~\cite{Gallegos16happier}and trajectory~\cite{González2008Track},to find the geographical content analysis with sentiment.
Associations are found between online sentiments and geographical factors, e.g happy regions are more likely to connect with each other~\cite{Alshamsi15milan},a high check-in density region usually presents a more positive mode~\cite{Gallegos16happier}, the whole process and development of a organized movement could be tracked on the social media~\cite{Alvarez2015Movement} and so on.

%and emotion
%complete survey
However, most existing work of this line employ simple statistical analysis to uncover the associations.
Such a coarse-grained analysis is distorted by latent variables, such as activity of the region.
Our work is the first to incorporate activity to obtain a fine-grained analysis.

\section{System Overview}
The section make an overview introduction for our post hoc framework to explain the functional region identification.
The whole structure is illustrated in Figure ~\ref{overview} while the corresponding steps as follow:

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.3]{overview.png}
  \label{overview}
  \caption{The overview structure of our framework}
\end{figure}

\emph{1.Functional region identification:}

The input data of first step is taxicabs OD pairs and POIs with coordinates.
We can get an activity-region matrix with value of the probability that each single region units belong to each activity by DMR.
And for the region, an activity with highest probability can be output as the function of this region.

\emph{2.Urban feature extraction:}

This step based upon the comments of different shops and activity-region matrix output in last step.
We select nouns with high frequency and low information entropy as our urban features.
Meanwhile, the sentiment analyse between urban feature and different activity is also illustrated in an activity-feature matrix.

\emph{3.Train Matrix Factorization model:}

With the certain activity-region matrix $\bm{r}$ and activity-feature matrix $\bm{u}$, we performed an matrix factorization model in form of Equation \ref{model} to get the region-feature matrix $\bm{v}$.
Training data comes from the given result of step 1 and step 2.

\emph{4.Output label explanation:}

The region-feature matrix $\bm{v}$ generated in step 3 in the input of this step
For each single region units,we can pick the urban features with highest sentiment score in $\bm{v}$.
These urban feature is the explanation for the functional region identification.

\section{Urban Feature}
\subsection{Functional Region Identification}
A city was segmented into single region units, which is regarded as documents with various themes in topic modeling.
DMR~\cite{Mimno2012DMR} was applied in functional region identification as explained in~\cite{Yuan2012FunctionRegion}.
DMR, whose structure is illustrated in Figure~\ref{DMR_ori}, is a topic clustering model based on Latent Dirichlet Allocation(LDA).
It input some metadata, e.g. authors and publishers, in addtion to basic LDA, which make the topic modeling more accurate.
As describe in~\cite{Yuan2012FunctionRegion},we regard flow increment calculated in regions as vocabulary in documents, which is marked as $m_{r,n}$ in Figure~\ref{DMR_ori}.
And POI located in different region units are correlate to the matadata of document, which is marked as $x_r$ in Figure~\ref{DMR_ori}.
\begin{figure}[h]
  \centering
  \includegraphics[scale=0.4]{DMR.png}
  \caption{The structure of DMR}
  \label{DMR_ori}
\end{figure}

\subsection{Urban Feature Extraction}
To get proper urban features for explaining different functions, we thought that urban features should be nouns with high frequency and low information entropy.

Considering the property of urban features, we separate comment sentences into single words and select all nouns of it.
With the help of coordinates, these words can correspond to different single units.
Take the sparseness of latent factor into account, we thought the words had better appear in more region units.
And we filer the high-frequency nouns as features.
If the frequency of a word is high enough, i.e. the shops commented with this word is much than a threshold $t$, we thought its frequency is high.
We set the threshold $t$ as 500.

Meanwhile, to make the urban features more distinguishing, we calculate the information entropy of words for different functional regions generated in Section 4.1.
Information entropy can measure the amount of information that a word contains.
Stronger distinguishing ability of a word takes lower information entropy.
The information entropy is defined as:
$$H[x_i,word_j] = -\sum_i p(x_i,word_j)\log p(x_i,word_j)$$
where $x_i$ stands for $i$th functional regions;
$p(x_i,word_j)$ is the proportion of shops located in $i$th functional regions commented with $j$th word in all shops located in $i$th functional regions.
We select features with lowest information entropy as our urban features, which are shown in Table \ref{urbanfeatures}.

\begin{table}[h]
\centering
\caption{Selected Urban Features}
\label{urbanfeatures}
\begin{tabular}{c|c|c|c|c}
\hline
features &  Alley & Brand & Center & Square\\
\hline
information entropy &  &  &  & \\
\hline
\end{tabular}
\end{table}

\subsection{Urban Feature Matrix}

\section{Explanation Model}
In this section,we describe our post hoc framework for explanation in detail.

\subsection{Latent Factor Model}
A classical model has Intuitive explanation is latent factor model.
It has a wide application in various fields.
The form of latent factor applied in recommendation system to predict rating of user $u$ to product $p$ as follow:
\begin{equation}
\hat{r}_{u,p}= g+b_u+b_p+\bm{x_u}^T\bm{y_p} \label{basic}
\end{equation}
where $g$ is a base rating in the system while $b_u$ and $b_p$ is the bias of user $u$ and product $p$;
$\bm{x_u}$ and $\bm{y_p}$ denote vectors of latent factors for corresponding user and product.

The matrix $r$ is rating that user $u$ give to product $p$.
Correlate users as activities, products as locations, we translated the matrix into the sentiment analysis score that actitivy $a$ to region $l$.
We can make our matrix based upon this stand latent factor model presented in Equation \ref{basic}.

\subsection{Our model}
Considering the temporal attribute of datesets, ~\cite{Zhang2017Model} has proposed an Historical Influence Aware Latent Factor(HIALF) model take the historical influence into account, which fit our intuition well and make the power of explanation stronger.

Therefore, we define our matrix factorization model as:
\begin{equation}
\hat{r}_{l,i,a}=b_l+U_a^TV_l+\beta(e_{a,i})\label{model}
\end{equation}
where $\hat{r}_{l,i,a}$ stands for the predicted probability at $i$-th time bin that region $l$ belongs to activity $a$;
$b_l$ is the bias of region $l$;
$\bm{U_a}$ and $\bm{V_l}$ represent the latent feature of activity $a$ and region $l$;
$e_{a,i}$ generated by prior expectation;
and $\beta(\cdot)$ demonstrates the bias curve changing with $e_{a,i}$.
Now we describe how to learn the $\beta(x)$ and generate a more realistic prior expectaion $e_{a,i}$.

\textbf{Modeling the bias curve $\beta(x)$}
The form of $\beta(x)$ is unknown.
We can constrain it with a data-driven approach.
Kernel regression is a kind of typical non-parameter learning and fit the $\beta(x)$ model well.

If we have a set of independent variables $e_l$ and dependent variables $v_l$, and $v_l=\beta(e_l)+\epsilon_l$ where $\epsilon_l$ is the noise from the standard normal distribution, we can defind $\beta(x)$ as:
$$\beta(x) = \frac{\sum_{k=1}^{n}w(x,e_l)v_l}{\sum_{k=1}^{n}w(x,e_l)}$$
where $w(x, x_i) = \exp(-\kappa(x-x_i)^2)$ and $\kappa$ is given by 10.
Actually, $e_l$ is given while $v_l$ is unknown.
Therefore, we set $e_l$ as several arithmetic progression within data range and $v_l$ is the corresponding unknown parameters so that we can learn them from datasets.

\textbf{Modeling the prior expectation $e_{a,i}$}
$$e_{a,i} = \frac{\sum_{k=1}^{i-1}\xi(i-k) r_{a,k}}{\sum_{k=1}^{i-1}\xi(i-k)}$$
where $\xi(d)=\exp(-\gamma\times d)$ is an exponential triggering kernel that models the decrease of histroy influence;
$\gamma$ controls the degree that history probability influence current probability;
and $r_{a,k}$ denotes the sentiment of activity $a$ when time bin $k$.

\subsection{Model Inference}
After build a matrix factorization model, we set up objective function as follow:
\begin{equation}
F=\sum\limits_{(p,l,u)\in\mathcal{M}}(r_{l,i,a}-\hat{r}_{l,i,a})^2+\lambda_1(b_l^2+||{\bm{y_r}}||_2^2)+\lambda_2(\sum\limits_{l}v_l^2)+\lambda_3(\sum\limits_{\alpha}\alpha_u^2)
\label{inference}
\end{equation}
where $\Theta$ stands for the unknown parameters, $\bm{y_p}$, $b_u$, $\alpha$ and $v_l$.
$\mathcal{M}$ contains all $(p,l,u)$ pairs, and each pair of it stands for the related parameter of region $l$ belongs to activity $a$ at time bin $i$.
$r_{r,i,a}$ is the probability generated by DMR while $\hat{r}_{r,i,a}$ is the predicted probability by Equation \ref{model}.
The first item of Equation \ref{inference} is quadratic sums of the difference between real probability $r_{r,i,a}$ and predected probability $\hat{r}_{r,i,a}$.
To prevent from overfitting, we three regularization terms and $\lambda_1$, $\lambda_2$ and $\lambda_3$ are given as hyperparameters.

To get our optimal $\bm{y_r}$, we aim to solve the following optimization problem and make the predicted probability $\hat{r}_{r,i,a}$ as close as possible to the real probability$r_{r,i,a}$:
$$\min\limits_{\Theta} F$$
The approach to get it is stochastic gradient descent(SGD) algorithm due to its efficiency in learning the parameter of objective function

\section{Experiments}
\subsection{Data Set}
The data set for our experiments including both mobility data and content data.
The mobility data including the pick-up and drop-off timestamps and coordinates in November of the year 2016 provided by DiDi, the biggest taxi platform in China.It contributed to the movement pattern of human mobility.
And the online content is crawled from a website with many comments similar to Yelp called DazhongDianping, which helps sentiment analysis of the regions.
\begin{table}[h]
\caption{Statistics of datasets}\label{dataset}
\begin{tabular}{|l|l|c|}
\hline
Datasets &  attributes & Value\\
\hline
comments & shops with coordinates(POI) & 109686\\
 & shops with comments & 50853\\
 & total comments & 3213264\\
\hline
taxicabs & effective orders & 7065937\\
 & effective days & 30 \\
\hline
road networks & geographical scope & [103.93$^\circ$E,104.21$^\circ$E] and [30.56$^\circ$N,30.79$^\circ$N]\\
 & road segments & 3712\\
 & percentage of major roads & 54.9\%\\
 & segmented regions & 901\\
\hline
\end{tabular}
\end{table}

\subsection{Preparation}
To form the basic region of city,we segment the urban area of city into region units by the major road network and make a map simplification.
The longitude of map range is [103.93,104.21] and latitude range is [30.56,30.79], which covers the main area of a city.
Raster-based model is more computationally efficient and succinct for territorial analysis, which is suitable for our map scenario.
We downloaded the major road network of this region\footnote{http://www.bigemap.com/} and rasterized the area into a 2000 $\times$ 2400 grid.
In the grid, the road network is converted to a binary image, as 1 stands for the road while 0 stands for the blank areas.

The main road data is present in Fig.\ref{raw_map}, including motorway, trunk, primary, secondary, tertiary and their links.
But the Fig.\ref{raw_map} is full of some unnecessary details, such as the lanes of a road and the overpasses, which disturb the distribution of regions.
As explained in~\cite{Yuan2015FunctionRegion}, the dilation and thinning process illustrated in Figure~\ref{dilation and thinning} are operated on the original road data to remove some small regions and simplify the map.
\begin{figure}[h]
    \centering
    \subfigure[Original segment of city]{
    \label{raw_map} %% label for first subfigure
    \includegraphics[width=1.5in]{chengdu_1.png}}
    \subfigure[After dilation operator]{
    \label{dilation} %% label for second subfigure
    \includegraphics[width=1.5in]{chengdu_2.png}}
    \subfigure[After thinning operator]{
    \label{thinning} %% label for second subfigure
    \includegraphics[width=1.5in]{chengdu_3.png}}
    \caption{The preparation process of road network}
    \label{dilation and thinning} %% label for entire figure
\end{figure}

We use DMR to identify regions into 8/4 types with different regions, which is illustrated in Figure~\ref{DMR_result}
The main function regions in the city is residence region, business region, study and science region, scenery regions and so on.

\begin{figure}[h]
\centering
\subfigure[After dilation operator]{
\label{thin_result} %% label for second subfigure
\includegraphics[width=2.3in]{FunctionalRegions_thinning.png}}
\subfigure[segmented region units of the city]{
\label{DMR_result} %% label for second subfigure
\includegraphics[width=2.3in]{FunctionalRegions_DMR.png}}
\caption{results of DMR}
\label{The results of functional region identification}
\end{figure}

\subsection{Evaluation Metrics}
We evaluate the functional region through two approaches. One is the percentage of regions that could find explanation label by our framework; while the other is the quality and accuracy of the explanation given by our framework.

For the evaluation of percentage of explainable region, we put forward the $explanation percentage$ to measure the proportion that regions with explainable labels as:
$$explanation\ percentage=\frac{|regions\ with\ explanation\ label|}{|all\ regions|}$$

For the second approach,we take a metric to measure the quality and accuracy of our explanation label.

\subsection{Baseline and other comparison}

\subsection{result}
\begin{table}[h]
\centering
\caption{Functional Regions and corresponding urban feature}\label{dataset}
\begin{tabular}{c|c}
\hline
Functions & Urban Feature\\
\hline
Business & \\
\hline
Residence & \\
\hline
Study & \\
\hline
Scenery & \\
\hline
\end{tabular}
\end{table}

\section{Conclusion}
Function regions identification is an important part of urban computing.
But its evaluation depend on human intuition and urban planning, which is hard to display in terms of statistics.
To make up for the lack of persuasive explanation, we proposed a post hoc framework to give persuasive explanation for functional regions identification in this paper.
Our datasets including over 3 million comments of shops and taxicabs OD pair trajectory in November 2016 generated by more than 7 million orders.
We utilized the framework to general most relative labels for every single region units as explanation.
According to the experiments performed in the datasets, our framework give an strong evidence to the identification and enhance its persuasiveness.
The result can help users and urban system designers easily recognize the region functions, which is helpful in a variety of urban applications, such as urban planning, location choosing for a business advertisement casting, and so on.

There are some directions can improve in the future work.
First is to full utilize our data sets of trajectory. 
Our trajectory datasets not only have origin and destination, but also include the detail points that users have passed by.
There are also some interesting patterns within these process points, and we can find new moving patterns in it. 
Second, we want to change our phrase-level explanation into sentence-level, which is more similar to natural language.

\begin{thebibliography}{25}
\bibitem{Zheng2014UrbanConcepts}Zheng Y, Capra L, Wolfson O, et al. Urban Computing:Concepts, Methodologies, and Applications[J]. Acm Transactions on Intelligent Systems \& Technology, 2014, 5(3):1-55.
    %综述城市计算的概念和框架
\bibitem{Zheng2011taxicabs}Zheng Y, Liu Y, Yuan J, et al. Urban computing with taxicabs[C]// International Conference on Ubiquitous Computing. ACM, 2011:89-98.
    %利用GPS轨迹检测出现有城市规划中存在的缺陷。结果包括：1）存在显著交通问题的区域对，2）连接结构以及它们之间的相关性。这些结果可以评估所实施的规划的有效性，例如一个城市新建的公路和地铁线路，并提醒城市规划者在构思未来规划时没有认识到的问题。
\bibitem{Meng2017Traffic}Meng, C.; Yi, X.; Su, L.; Gao, J.; Zheng, Y. City-wide Traffic Volume Inference with Loop Detector Data and Taxi Trajectories. Proceedings of the 25th ACM SIGSPATIAL International Conference on Advances in Geographic Information Systems,ACM,2017, 1:1-1:10
    %提出了一个框架来推断城市交通流量信息与环路探测器和出租车轨迹。

\bibitem{Vatsavai2011Remote}Vatsavai R R, Bright E, Varun C, et al. Machine learning approaches for high-resolution urban land cover classification:a comparative study[C]// International Conference and Exhibition on Computing for Geospatial Research \& Application, Com.geo 2011, Washington, Dc, Usa, May. DBLP, 2011:1-10.
    %利用遥感数据生成主题地图
\bibitem{Vatsavai2010Remote}Vatsavai R R, Cheriyadat A, Gleason S. Unsupervised Semantic Labeling Framework for Identification of Complex Facilities in High-Resolution Remote Sensing Images[C]// IEEE International Conference on Data Mining Workshops. IEEE Computer Society, 2010:273-280.
    %利用遥感数据生成主题地图

\bibitem{Ratti2010Telecom}Ratti C, Sobolevsky S, Calabrese F, et al. Redrawing the Map of Great Britain from a Network of Human Interactions[J]. Plos One, 2010, 5(12):e14248.
    %利用电信数据库来划分地图。
\bibitem{Yuan2010NextPassenger}Yuan J, Zheng Y, Zhang L, et al. Where to find my next passenger[C]// Proceedings of the 13th international conference on Ubiquitous computing. ACM, 2011:109-118.
\bibitem{Ge2011TaxiBusiness}Ge Y, Liu C, Xiong H, et al. A taxi business intelligence system[C]// ACM SIGKDD International Conference on Knowledge Discovery and Data Mining, San Diego, Ca, Usa, August. DBLP, 2011:735-738.
    %利用轨迹点来提升司机收入
\bibitem{Farmer2009Overview}Farmer, C. J. (2009). Data driven functional regions. In Proceedings of 10th International Conference on GeoComp
    %对之前的functional region做了一个总结。
\bibitem{Garg2018Route}Nandani Garg and Sayan Ranu. Route Recommendations for Idle Taxi Drivers: Find Me the Shortest Route to a Customer!. In KDD ’18: The 24th ACM SIGKDD International Conference on Knowledge Discovery \& Data Mining, August 19–23, 2018, London, United Kingdom. 2018.
    %通过Minimizing Distance through Monte Carlo Tree Search来对出租车司机进行路线推荐，使其提高效率，降低空载率。
\bibitem{Goddard1970EarliestFunction}Goddard J B. Functional Regions within the City Centre: A Study by Factor Analysis of Taxi Flows in Central London[J]. Transactions of the Institute of British Geographers, 1970, 49(49):161-182.
    %最早提出用出租车轨迹来做功能区划分。但是早期数据都来自司机填写调查问卷，不够准确且数据较小，因此研究的东西比较初步，只能表明伦敦中心区包含许多具有强大内部联系的非常独立的功能区。
\bibitem{Karlsson2006FunctionalRegionSummary}Karlsson C, Olsson M. The identification of functional regions: theory, methods, and applications[J]. Annals of Regional Science, 2006, 40(1):1-18.
    %对现有（2006）的功能区识别方法做了总结，将方法归纳为三类：
    %（1）基于不同城市核心区之间通勤频率的比较的当地劳动力市场方法
    %（2）通勤区方法（Killian等。1993年，它建立在现有的城市相互依存的基础上，并衡量地区之间双向通勤联系的强度
    %(3)无障碍方法，它利用两种无障碍方式：雇主与工人的接触。获得工作机会。
\bibitem{Newman2004ModularityFunction}Newman M E J. Detecting community structure in networks[J]. European Physical Journal B, 2004, 38(2):321-330.
    %模块化方法的鼻祖
\bibitem{Ball1980Definition}R.M. Ball. The use and definition of Travel-to-Work Areas in Great Britain: Some problems[J]. Regional Studies, 1980, 14(2):125-139.
    %早期的Urban function定义
\bibitem{Yuan2012FunctionRegion}Yuan J, Zheng Y, Xie X. Discovering regions of different functions in a city using human mobility and POIs[C]// ACM SIGKDD International Conference on Knowledge Discovery and Data Mining. ACM, 2012:186-194.
    %套用LDA，根据功能划分城市区域
\bibitem{Yuan2015FunctionRegion}Yuan N J, Zheng Y, Xie X, et al. Discovering Urban Functional Zones Using Latent Activity Trajectories[J]. Knowledge \& Data Engineering IEEE Transactions on, 2015, 27(3):712-725.
    %引入latent activity，对他自己2012的论文做了一个拓展，提高精确度
\bibitem{Yuan2018FunctionRegion}Yuan N J, Zheng Y, Xie X. Discovering Functional Zones in a City Using Human Movements and Points of Interest[J]. 2018.
    %对之前functional region工作的一个总结，没有再加入新的提高
\bibitem{Zhang2015Fuel}Zhang F, Yuan N J, Wilkie D, et al. Sensing the Pulse of Urban Refueling Behavior:A Perspective from Taxi Mobility[J]. Acm Transactions on Intelligent Systems \& Technology, 2015, 6(3):1-23.
    %提出了一个智能系统，用于利用出租车轨迹、POI、道路网络和在线用户的报告来发现城市加油行为。
\bibitem{Shang2014Pollution}Shang J, Zheng Y, Tong W, et al. Inferring gas consumption and pollution emission of vehicles throughout a city[J]. 2014:1027-1036.
    %本文利用车辆轨迹（例如出租车）的GPS轨迹，即时推断出在城市时速路段上行驶的车辆的气体消耗量和污染排放量。
\bibitem{Wu2017Reachable}Wu G, Ding Y, Li Y, et al. Mining Spatio-Temporal Reachable Regions over Massive Trajectory Data[C]// IEEE, International Conference on Data Engineering. IEEE, 2017:1283-1294.
    %寻找时空可达区域
    %
\bibitem{Wang2018Deep}D Wang,J Zhang,W Cao,J Li,Y Zheng.When Will You Arrive? Estimating Travel Time Based on Deep Neural Networks.In International AAAI Conference on Web and Social Media, 2011.
    %用深度神经网络来预测旅行时间
    %
\bibitem{Alvarez2015Movement}R. Alvarez, D. Garcia, Y. Moreno, and F. Schweitzer. Sentiment cascades in the 15m movement. EPJ Data Science, 4(1):1–13, 2015
    %从社交网站上挖掘15M这个运动的组织、形成与爆发
    %
\bibitem{Alshamsi15milan}Alshamsi A, Awad E, Almehrezi M, et al. Misery loves company: happiness and communication in the city[J]. Epj Data Science, 2015, 4(1):7.
    %把米兰城区划分成100*100区域，根据社交网络发现情绪积极的region更喜欢和其他积极的region交流。
\bibitem{Cheng2011Check-in}Z. Cheng, J. Caverlee, K. Lee, and D. Sui. Exploring millions of footprints in location sharing services. In International AAAI Conference on Web and Social Media, 2011.
    %check-in数据来自整个美国.         用户有周期性的行为模式（weekly），受地理、经济、个人社会地位（粉丝越多地位越高）约束，需要content来更好地理解.引入回转半径来帮助sentiment（实际文本分析还是靠字典）
\bibitem{González2008Track}González M C, Hidalgo C A. Understanding individual human mobility patterns[J]. Nature, 2008, 453(7196):779-782.
    %追踪100000名手机用户6个月，人体轨迹表现出高度的时间和空间规律。说是human mobility, 但是数据集是长期轨迹包含很多点
\bibitem{Gallegos16happier}Gallegos L, Huang A, Huang A, et al. Geography of Emotion: Where in a City are People Happier?[C]// International Conference Companion on World Wide Web. International World Wide Web Conferences Steering Committee, 2016:569-574.
    %将其他人的统计方法用于划分区域的签到数据集中，比较得出结论：① 签到数据越多的地方积极性越高，消极性越低。②发现在签到数据多的地方签到的用户，活动范围都比较大。可以理解为这个地方比较吸引人，人们愿意走更远的路去这里玩。③结合人口普查的结果，发现签到数据多的地方所居住的居民年龄更大、非西班牙裔人口更多、受教育程度更高
\bibitem{Zhao2016POI}Zhao S, King I, Lyu M R. A Survey of Point-of-interest Recommendation in Location-based Social Networks[J]. 2016.
    %对截止16年最新的50篇与LSBN和POI相关的顶会论文做了一个总结，不知道能不能用上
\bibitem{Zhang2018SIGIR}Yongfeng Zhang, Yi Zhang, and Min Zhang. SIGIR 2018 Workshop on ExplainAble Recommendation and Search (EARS 2018). In Proceedings of The 41st International ACM SIGIR Conference on Research and Development in Information Retrieval (SIGIR ’18). 2018.
    %概述了推荐系统中explanation的重要性
\bibitem{Zhang2018Survey} Zhang Y, Chen X. Explainable Recommendation: A Survey and New Perspectives[J]. arXiv preprint arXiv:1804.11192. 2018.
    %介绍了推荐系统解释的调查概述和未来新视角。主要看了explanation的几种形式
    %基于内容、特征的explanation：利用用户a感兴趣的特征和商品b所拥有的特征来匹配。用户对推荐系统的满意度和信任度与解释高度相关。形式有雷达图等。
    %基于文本的句子形式explanation：在评论等数据中获得更加完善的用户倾向，生成文本句子作为解释。情感抽取工具Sentires（toolkit是一个jre，已得到，未运行），可抽取feature-Opinion-sentiment三元组
\bibitem{Herlocker2000Explanation} Herlocker J L, Konstan J A, Riedl J. Explaining collaborative filtering recommendations[C]// Acm Conference on Computer Supported Cooperative Work. 2000:241-250.
    %为了证明explanation facilities 能提高推荐系统的acceptance，本文通过对210个MovieLens用户的调查，发现绝大部分用户希望在推荐结果中加入explanation。
    %
\bibitem{Ferwerda2012ContentCorrelate} Ferwerda B, Swelsen K, Yang E. Explaining Content-Based Recommendations[J]. Bruceferwerda Com.
    %这篇只发表在作者自己的网站上，但是有被引用。内容是选择过多时，用户满意度会下降。这篇文章通过实验证明，用户的满意度和解释性有关。
    %
\bibitem{Zhang2015Sentires}  Zhang Y, Zhang M, Liu Y, et al. Boost Phrase-level Polarity Labelling with Review-level Sentiment Classification[J]. Computer Science, 2015.
    Zhang Y, Lai G, Zhang M, et al. Explicit factor models for explainable recommendation based on phrase-level sentiment analysis[C]// International Acm Sigir Conference on Research \& Development in Information Retrieval. ACM, 2014:83-92.
    %开发了一个情感抽取工具Sentires，在短语这个级别层面上做情感极性标注。可以从某个产品领域的产品评论中抽取“feature-oppion-sentiment”这样的三元组。例如，在手机领域，“noise-high-negative”，“screen-clear-positive”等。
    %对于我们的大众点评数据，也可以“feature-opinion-sentiment”。其中feature表示52个标签，opinion表示评分，sentiment表示根据评分是积极or消极（与feature其实无关）
\bibitem{Costa2017LSTM} Costa F, Ouyang S, Dolog P, et al. Automatic Generation of Natural Language Explanations[J]. 2017.
    %提出了一个LSTM模型来生成评论。并提出八个评估生成文字可读性的测量标准
\bibitem{Preece2018ExplainAI}Preece A. Asking 'Why' inAI: Explainability of intelligent systems - perspectivesand challenges. Intell Sys Acc Fin Mgmt. 2018;25:63–72.
    %explain AI
\bibitem{Korpan2017navigation}Korpan, R., Epstein, S.L., Aroor, A., Dekel, G.: WHY: Natural explanations from a robot navigator. In: AAAI 2017 Fall Symposium on Natural Communication for Human-Robot Collaboration.
\bibitem{Jose2018cognitive}Jose M. Alonso, Corrado Mencar. Building Cognitive Cities with Explainable Artificial Intelligent Systems. In Proceedings of the First International Workshop on Comprehensibility and Explanation in AI \& ML, 2017.
\bibitem{Lu2011Label}Lu Y, Castellanos M, Dayal U, et al. Automatic construction of a context-aware sentiment lexicon:an optimization approach[C]// International Conference on World Wide Web, WWW 2011, Hyderabad, India, March 28 - April. DBLP, 2011:347-356.
\bibitem{Zhang2017Model}Zhang X , Zhao J , Lui J C S . Modeling the Assimilation-Contrast Effects in Online Product Rating Systems: Debiasing and Recommendations[C]// Eleventh Acm Conference on Recommender Systems. ACM, 2017.
\bibitem{Mimno2012DMR}Mimno D, Mccallum A. Topic Models Conditioned on Arbitrary Features with Dirichlet-multinomial Regression[J]. University of Massachusetts - Amherst, 2012, 2008:411--418.
\end{thebibliography}
\end{document}